---
title: "Bus Ridership Analysis"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

#### Create a vector s1 whose elements correspond to the number of passengers who can possibly get on board the bus at a particular stop
```{r}
s1 = c(0, 1, 2)
```

#### Create a vector p1 whose elements correspond to the respective chances of the number of passengers identified in the previous question getting on board the bus at a particular stop.
```{r}
p1 = c(0.5, 0.4, 0.1)
```

#### Create a 2x10-matrix P such that the jth column of this matrix should have a passenger on the bus’s continuing-on and getting-off chances for the jth stop for j = 1, . . . , 10.
```{r}
P = matrix(c(0.99, 0.01, 0.2, 0.8, 0.2, 0.8, 0.8, 0.2, 0.8, 0.2, 0.8, 0.2, 0.8, 0.2, 0.2, 0.8, 0.2, 0.8, 0.0, 1.0), nrow = 2, ncol = 10)
P
```

#### Create a vector s2 whose elements correspond to the encoded versions of a passenger staying
on the bus and getting off of the bus at a particular stop, respectively.
```{r}
s2 = c(0, 1)
```

####

```{r}
n = 30 # zero passengers in the bus at stop-1
for (j in c(1:9)){
  # Simulating the continuing on or getting off of the bus
  if (n > 0){ # if there is at least one passenger in the bus
    # Update the current number of passengers in the bus by subtracting off the number of passengers that get off of the bus at the j-th stop
    n = n - sum(sample(s2, size = n, replace = T, prob = P[, j]))
  }
  # Update the current number of passengers in the bus by adding the number of passengers that board the bus at the j-th stop
  n = n + sample(s1, size = 1, prob = p1)
  cat(sprintf('No. of passenger after stop-%d = %d\n', j, n))
}

```


