---
title: "Error metrics"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
library(titanic)
library(caret)
library(pROC)
library(PRROC)
library(splitstackshape)
```


#### Load the in-built titanic_train dataset
```{r}
data("titanic_train")
df = titanic_train
head(df, 2)
#str(df)
```

#### Preprocess the dataset (note the chained dplyr usage)
```{r}
# 
df = titanic_train %>%
  select(Survived, Pclass, Sex, Age, SibSp, Parch, Fare) %>%
  mutate(Sex = as.factor(Sex),
         Survived = as.factor(Survived)) %>%
  na.omit()  # Remove rows with missing values for simplicity
head(df, 2)
str(df)
```

#### Explore the class distribution (imbalanced target: Survival, 1 means survived, 0 means did not survive)
```{r}
table(titanic_train$Survived)
# Graphical approach
dfLabel = setNames(data.frame(table(titanic_train$Survived)), c('Label', 'Frequency'))
p = ggplot(data = dfLabel) +
  geom_bar( aes(x = Frequency, y = Label),
            stat = 'identity', fill = 'steelblue', width = 0.3) +
  theme(text = element_text(size = 14, face = 'bold'),
  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  xlab('') + ylab('Count') +
  ggtitle('Distribution of Output Labels')
p
```

#### Stratified train-test split of the dataset
```{r}
trainIndex = as.numeric(rownames(as.data.frame(stratified(df, 'Survived', 0.8, set.seed(42)))))
X_train = df[trainIndex, ]
X_test = df[-trainIndex, ]
```


#### Fit a logistic regression model using the train data to predict the output class
```{r}
model <- glm(Survived ~ ., data = X_train, family = binomial)
summary(model)
```

#### Predict on the test data
```{r}
pred_prob = predict(model,newdata = X_test, type = "response")
# Convert probabilities to binary outcomes
threshold = 0.5
pred_class = ifelse(pred_prob>threshold,1,0)
```


#### Build the confusion matrix and interpret the error metrics
```{r}
confusionMatrix(as.factor(pred_class), as.factor(X_test$Survived))
```

#### Plot the ROC curve for different thresholds
```{r}
# ROC curve and AUC
roc_obj = roc(X_test$Survived, pred_prob)
auc_val = auc(roc_obj)

# Plot the ROC curve
plot(roc_obj, col = "blue", main = paste("ROC Curve (AUC =", round(auc_val, 2), ")"))

### Precision-Recall Curve ###
# Create PRROC object for Precision-Recall curve
pr = pr.curve(scores.class0 = pred_prob[X_test$Survived == 1], 
               scores.class1 = pred_prob[X_test$Survived == 0],
              curve = TRUE)

# Plot Precision-Recall curve
plot(pr, main = "Precision-Recall Curve")
```



