---
title: "APS Lab 3"
output: html_document
date: "2025-10-13"
---
```{r}
library(MASS)
library(e1071)
data(epil)
ndata = epil
```

Step 1
```{r}
Y = ndata[,1]
hist(Y, breaks=40,freq=TRUE)

```
Step 2
```{r}
breaks = seq(min(Y)-0.5, max(Y)+0.5, by=1)
hist(Y, breaks = breaks, freq=FALSE, main = "PMF of Y")
skewness(Y)
```
```{r}
my = mean(Y)
mv = var(Y)
print(my)
print(mv) #overdispersed data
```
Step 3:
```{r}
poismod = fitdistr(Y,densfun="poisson")
lam = poismod$estimate
sprintf("Log likelihoof:%f", poismod$loglik)
```
```{r}
k = 0:max(Y)
pois_pmf = dpois(k, lambda = lam)
hist(Y, breaks = breaks, freq=FALSE, main = "PMF of Y")
lines(k, pois_pmf, lwd=1, col="red")
```
Step 4: 
```{r}
nbmod = fitdistr(Y, densfun = "Negative Binomial")
nb_size = nbmod$estimate[1]
nb_mu = nbmod$estimate[2]
nb_pmf = dnbinom(k, nb_size, mu=nb_mu)
sprintf("Log likelihood: %f", nbmod$loglik)
hist(Y, breaks=breaks, freq = FALSE, main="PMF of Y")
lines(k, nb_pmf, lwd=1, col="red")
```
Step 5:
```{r}
nr = nrow(ndata)
s = sample(x=1:nr, size=0.8*nr, replace = FALSE)
traindata = ndata[s,]
testdata = ndata[-s,]
linearmod = lm(y~lbase+lage+trt, data=traindata)
ypredtrain = predict(linearmod, traindata)
ypredtest = predict(linearmod, testdata)
rmsetrain = sqrt(mean((traindata[,1]-ypredtrain)^2))
maetrain = mean(abs(traindata[,1]-ypredtrain))
rmsetest = sqrt(mean((testdata[,1]-ypredtest)^2))
maetest = mean(abs(testdata[,1]-ypredtest))
likely = logLik(linearmod)
sprintf("Train RMSE: %f",rmsetrain)
sprintf("Train MAE: %f",maetrain)
sprintf("Test RMSE: %f",rmsetest)
sprintf("Test MAE: %f",maetest)
sprintf("Log Likelihood: %f",likely)
```
Step 6:
```{r}
poistrgmod = glm(y~lbase+lage+trt, data=traindata, family="quasipoisson")
ypredtrain = predict(poistrgmod, traindata)
ypredtest = predict(poistrgmod, testdata)
rmsetrain = sqrt(mean((traindata[,1]-ypredtrain)^2))
maetrain = mean(abs(traindata[,1]-ypredtrain))
rmsetest = sqrt(mean((testdata[,1]-ypredtest)^2))
maetest = mean(abs(testdata[,1]-ypredtest))
likely = logLik(poistrgmod)
sprintf("Train RMSE: %f",rmsetrain)
sprintf("Train MAE: %f",maetrain)
sprintf("Test RMSE: %f",rmsetest)
sprintf("Test MAE: %f",maetest)
sprintf("Log Likelihood: %f",likely)
```
7. Develop a Generalized Linear Model (GLM) regression using negative binomial 
likelihood as the loss to model y vs f in 5. Evaluate MAE, RMSE and log likelihood 
values. Which model (5,6,7) is best (best in what metric?)?
```{r}
nbregmod = glm.nb(y~lbase+lage+trt, data=traindata)
ypredtrain = predict(nbregmod, traindata)
ypredtest = predict(nbregmod, testdata)
rmsetrain = sqrt(mean((traindata[,1]-ypredtrain)^2))
maetrain = mean(abs(traindata[,1]-ypredtrain))
rmsetest = sqrt(mean((testdata[,1]-ypredtest)^2))
maetest = mean(abs(testdata[,1]-ypredtest))
likely = logLik(nbregmod)
sprintf("Train RMSE: %f",rmsetrain)
sprintf("Train MAE: %f",maetrain)
sprintf("Test RMSE: %f",rmsetest)
sprintf("Test MAE: %f",maetest)
sprintf("Log Likelihood: %f",likely)
```
8. Exploratory: Perform a log transformation on y(z=ln(1+y)). Implement regression 
(z~lbase+lage+trt) using the methods in steps 5,6, and 7. Explain your observations.

```{r}
nr = nrow(ndata)
s = sample(x=1:nr, size=0.8*nr, replace = FALSE)
traindata = ndata[s,]
testdata = ndata[-s,]
linearmod = lm(log(1+y)~lbase+lage+trt, data=traindata)
zpredtrain = predict(linearmod, traindata)
zpredtest = predict(linearmod, testdata)
ypredtrain = exp(zpredtrain)-1
ypredtest = exp(zpredtest)-1
rmsetrain = sqrt(mean((traindata[,1]-ypredtrain)^2))
maetrain = mean(abs(traindata[,1]-ypredtrain))
rmsetest = sqrt(mean((testdata[,1]-ypredtest)^2))
maetest = mean(abs(testdata[,1]-ypredtest))
likely = logLik(linearmod)
sprintf("Train RMSE: %f",rmsetrain)
sprintf("Train MAE: %f",maetrain)
sprintf("Test RMSE: %f",rmsetest)
sprintf("Test MAE: %f",maetest)
sprintf("Log Likelihood: %f",likely)
```
```{r}
poistrgmod = glm(log1p(y)~lbase+lage+trt, data=traindata, family="quasipoisson")
zpredtrain = predict(poistrgmod, traindata)
zpredtest = predict(poistrgmod, testdata)
ypredtrain = exp(zpredtrain)-1
ypredtest = exp(zpredtest)-1
rmsetrain = sqrt(mean((traindata[,1]-ypredtrain)^2))
maetrain = mean(abs(traindata[,1]-ypredtrain))
rmsetest = sqrt(mean((testdata[,1]-ypredtest)^2))
maetest = mean(abs(testdata[,1]-ypredtest))
likely = logLik(poistrgmod)
sprintf("Train RMSE: %f",rmsetrain)
sprintf("Train MAE: %f",maetrain)
sprintf("Test RMSE: %f",rmsetest)
sprintf("Test MAE: %f",maetest)
sprintf("Log Likelihood: %f",likely)
```
```{r}
nbregmod = glm.nb(log1p(y)~lbase+lage+trt, data=traindata)
zpredtrain = predict(nbregmod, traindata)
zpredtest = pred                    ict(nbregmod, testdata)
ypredtrain = exp(zpredtrain)-1
ypredtest = exp(zpredtest)-1
rmsetrain = sqrt(mean((traindata[,1]-ypredtrain)^2))
maetrain = mean(abs(traindata[,1]-ypredtrain))
rmsetest = sqrt(mean((testdata[,1]-ypredtest)^2))
maetest = mean(abs(testdata[,1]-ypredtest))
likely = logLik(nbregmod)
sprintf("Train RMSE: %f",rmsetrain)
sprintf("Train MAE: %f",maetrain)
sprintf("Test RMSE: %f",rmsetest)
sprintf("Test MAE: %f",maetest)
sprintf("Log Likelihood: %f",likely)
```

