---
title: "Exp4"
output: html_document
date: "2025-10-27"
---

Generate a binomial random variable b~B(10,0.4), where n=10 and p=0.4. Repeat the 
experiment for 10000 trials and store it in a vector ‘x’.
```{r}
samples = numeric(10000)
samples = rbinom(10000, 10, 0.4)
#for (i in 1:10000){
#samples[i] = rbinom(1, 10, 0.4)  
#}
```

Convert x to a zero mean and unit variance random variable y. Draw a Density 
histogram for y. 
```{r}
y = scale(samples, center = TRUE, scale = TRUE)
hist(y, breaks = 40, freq = FALSE)
```
Using the values in y, overlay a normal density function on the histogram in step 2
```{r}
k = seq(min(y), max(y), by = 0.001)
gpdf = dnorm(k,mean(y), sd = sd(y))
hist(y, breaks = 40, freq = FALSE)
lines(k, gpdf, col = "red")

```
Repeat step 1-3 by taking n as 50, 100 and 500. What are your observations?
```{r}
n = c(50, 100, 500, 1000)
samples = numeric(10000)
for (nn in n){
  for(i in 1:10000){
  samples[i]=rbinom(1,nn,0.4)
}
y = scale(samples)
k = seq(min(y), max(y), by = 0.001)
gpdf = dnorm(k,mean(y), sd = sd(y))
hist(y, breaks = 40, freq = FALSE)
lines(k, gpdf, col = "red")
}

```
Download “biketental.csv”. Define a variable (Z) as number of rentals/day. Transform Z 
to a zero mean and unit variance variable (Y). Write a code to plot the Density 
histogram for Y. In Density histogram x-axis is the variable and y-axis is the density 
(count/Total count). Choose number of bins as 40

```{r}
ndata = read.csv("E:/251100670036/APS/bikerental.csv")
Z = ndata[,13]
Y = scale(Z)
hist(Y, breaks = 40, freq=FALSE)
```
Using the values in Y, overlay a normal density function on the histogram in step 5.
```{r}
k = seq(min(Y), max(y), by = 0.001)
gpdf = dnorm(k, mean = mean(Y), sd = sd(Y))
hist(Y, breaks = 40, freq=FALSE)
lines(k, gpdf, col = "red")
```

Using correlation as metric, identify two features (out of columns 1-11) which have 
maximum correlation with Y.

```{r}
corr_metric = numeric(length = ncol(Y))
X = ndata[,2:12]
X= scale(X)
for(i in 1:11){
    corr_metric[i] = abs(cor(X[,i],Y))
}
corr_metric = order(corr_metric, decreasing = TRUE)[1:2]
corr_metric
```

Implement a linear regression (Y vs (Two features in step 7)). Split the dataset into 
training and test set. Find mean absolute error for the training and test set.
```{r}
nr = nrow(X)
Xnew = X[,corr_metric]
Xnew = cbind(Y, Xnew)
s = sample(1:nr, size = 0.8*nr, replace = FALSE)
train = as.data.frame(Xnew[s,])
test = as.data.frame(Xnew[-s,])
lmod = lm(V1~., data = train)
ytrainPred = predict(lmod, train)
ytestPred = predict(lmod, test)
trainmae = mean(abs(ytrainPred-train[,1]))
testmae = mean(abs(ytestPred-test[,1]))
trainmae
testmae
```
Plot a histogram for the variable e=Y-Yp, where Yp are the predictions on the training 
set instep 9. Fit a Gaussian distribution to e and show it on the histogram.

```{r}
e = Y - ypredtrain
k = seq(min(e), max(e), by = 0.001)
gaus_pdf = dnorm(k, mean = mean(e), sd = sd(e))
hist(e, breaks = 40, freq = FALSE)
lines(k, gaus_pdf, col = "red" )
```

Exploratory: Implement linear regression in step 8 for all combinations of 2 features 
(55 combinations). Print Train MAE and Test MAE in each case.

```{r}
count = 0
for(i in 2:12){
  for (j in i:12){
    if (i==j){
      next
    }
    Xnew = cbind(ndata[,i], ndata[,j])
    Xnew = cbind(Y=Y, Xnew)
    s2 = sample(1:nr, size = 0.8*nr, replace = FALSE)
    train = as.data.frame(Xnew[s2,])
    test = as.data.frame(Xnew[-s2,])
    lmod = lm(V1~., data = train)
    ytrainPred = predict(lmod, train)
    ytestPred = predict(lmod, test)
    trainmae = mean(abs(ytrainPred-train[,1]))
    testmae = mean(abs(ytestPred-test[,1]))
    print(paste("For Column: ", i, " and ", j, "trainMAE: ", trainmae, "and TestMae: ", testmae))
    count=count+1
  }
}
print(count)


```
























