---
title: "Experiment_5"
output: html_document
date: "2025-11-03"
---

1. Generate nsamples=100 samples from uniform distribution (X~U(0,5)). Find sample 
mean of these 100 samples. Repeat the experiment for ntrials=10000 trials and store 
sample means in a vector ‘mx’. Plot density histogram of mx and overlay a Gaussian 
curve.
```{r}
library(moments)
ntrials = 10000
nsamples = 100 #whether n is large or not
mx = vector("numeric", ntrials )
mx = replicate(ntrials, mean(runif(nsamples, 0, 5)))

k = seq(min(mx), max(mx), by = 0.001)
gaus_pdf = dnorm(k, mean(mx), sd(mx))
hist(mx, breaks = 40, freq = FALSE)
lines(k, gaus_pdf, col = "red")
sprintf("Skewness: %f ", skewness(mx))
sprintf("Kurtosis: %f", kurtosis(mx))

#since skewness is close to 0, the random variable is gaussian
```
2. Generate nsamples=100 samples from Exponential distribution (X~Exp(5)). Find 
sample mean of these 100 samples. Repeat the experiment for ntrials=10000 trials 
and store sample means in a vector ‘mx’. Plot density histogram of mx and overlay a 
Gaussian curve.
```{r}
nsamples = 200
mx = replicate(ntrials, mean(rexp(nsamples,5)))
k = seq(min(mx), max(mx), by = 0.001)
gaus_pdf = dnorm(k, mean(mx), sd(mx))
hist(mx, breaks = 40, freq = FALSE)
lines(k, gaus_pdf, col = "red")
sprintf("Skewness: %f ", skewness(mx))
sprintf("Kurtosis: %f", kurtosis(mx))
#since the skewness is closer to 0, it is gaussian.

```
3. Generate nsamples=100 samples from Gaussian distribution (X~N(0,5)). Find sample 
mean of these 100 samples. Repeat the experiment for ntrials=10000 trials and store 
sample means in a vector ‘mx’. Plot density histogram of mx and overlay a Gaussian 
curve.

```{r}
nsamples = 100
mx = replicate(ntrials, mean(rnorm(nsamples, 0, 5)))
k = seq(min(mx), max(mx), by = 0.001)
gaus_pdf = dnorm(k, mean(mx), sd(mx))
hist(mx, breaks = 40, freq = FALSE)
lines(k, gaus_pdf, col = "red")
sprintf("Skewness: %f ", skewness(mx))
sprintf("Kurtosis: %f", kurtosis(mx))
#Skewness is closer to 0, hence the random vairable is gaussian
```

4. Generate nsamples=100 samples. Each sample can randomly come from one of the 
following three distributions (X~U(0,5), X~Exp(5), X~N(0,5)). Find sample mean of 
these 100 samples. Repeat the experiment for ntrials=10000 trials and store sample 
means in a vector ‘mx’. Plot density histogram of mx and overlay a Gaussian curve.

```{r}
library(moments)

ntrials = 10000
nsamples = 10
mx = numeric(ntrials)

suppressWarnings({
  for (i in 1:ntrials) {
    x = numeric(nsamples)
    for (j in 1:nsamples) {
      s = sample(1:3, size = 1)
      if (s == 1) {
        x[j] = runif(1, 0, 5)       # generate ONE uniform sample
      } else if (s == 2) {
        x[j] = rexp(1, rate = 5)    # generate ONE exponential sample
      } else {
        x[j] = rnorm(1, 0, 5)       # generate ONE normal sample
      }
    }
    mx[i] = mean(x)
  }
})

# Create Gaussian fit values
k = seq(min(mx), max(mx), length.out = 1000)
gaus_pdf = dnorm(k, mean(mx), sd(mx))

# Plot histogram with Gaussian overlay
hist(mx, breaks = 40, freq = FALSE)
lines(k, gaus_pdf, col = "red", lwd = 2)

# Print skewness and kurtosis
cat(sprintf("Skewness: %f\n", skewness(mx)))
cat(sprintf("Kurtosis: %f\n", kurtosis(mx)))

#Skewness is closer to 0, hence the random vairable is gaussian
```
5.  In each step 1-4, decrease the variable ‘nsamples’ and identify the value 
when the mean distribution stops being Gaussian (i.e. identify how much ‘nsamples’ 
should be considered as large).

For the uniform distribution, the mean distribution is not gaussian at nsamples = 1 but it is gaussian for nsamples = 2. Hence 2 can be considered as large.

For Exponential distribution, the mean distribution is skewed for nsamples <40. Hence nsamples can be considered large for nsamples values >=40

For Gaussian distribution, any value > 0 is will be gaussian. 

For the combined samples, the skewness is least at nsamples = 7. Hence large value can be considered for nsamples >=8

6. Generate simulated data: Select x in the range (1,5,by=0.01); y=a+b*x+d where a=1.2, 
b=2.1, d~N(0,2). Plot a scatter plot y vs x

```{r}
a=1.2
b = 2.1
x = seq(1, 5, by = 0.01)
y = numeric(length(x))
for (j in 1: length(x)){
  y[j] = a+b*x+rnorm(1, mean=0, sd =2)
}
plot(x,y)
```

7. Train a linear model on y vs x. Find the coefficients a, b and mean of error residuals 
(me)
```{r}
train = as.data.frame(cbind(y,x))
mymod = lm(y~x, data = train)
ypred = predict(mymod, train)
me = mean(y-ypred)
ea = mymod$coefficients[1]
eb = mymod$coefficients[2]

```

8. Repeat steps 6-7 for 1000 trials. Evaluate a, b and me for each trial and store it in a 
vector. Plot three density histogram (a) a, (b) b, and (c) me. Overlay a Gaussian curve in 
each case. Use the plot, skewness and Kurtosis values to identify if these variables are 
from Gaussian distribution.

```{r}
ntrials=1000
v1=numeric(ntrials)
v2=numeric(ntrials)
v3=numeric(ntrials)
for(i in 1:ntrials)
{
  a=1.2
  b=2.1
  x=seq(1,5,by=0.01)
  y=numeric(length(x))
  for(j in 1:length(x))
  {
    y[j]=a+b*x[j]+rnorm(1,mean=0,sd=2)#y=a+b(x)+d
  }
  #learn alinear regression model
  train=as.data.frame(cbind(x,y))
  mymod=lm(y~x,data=train)
  ypred=predict(mymod,train)
  v1[i]=mean(y-ypred) #variable 1
  v2[i]=mymod$coefficients[1] #variable 2
  v3[i]=mymod$coefficients[2] #variable 3
}
#for ME
hist(v1,breaks = 40,freq = FALSE)
k = seq(min(v1), max(v1), by = 1e-17)
gpdf = dnorm(k, mean(v1), sd(v1))
lines(k, gpdf, col = "red")
cat(sprintf("Skewness: %f\n", skewness(v1)))
cat(sprintf("Kurtosis: %f\n", kurtosis(v1)))

```
```{r}
#for A
hist(v2,breaks = 40,freq = FALSE)
k = seq(min(v2), max(v2), by = 0.001)
gpdf = dnorm(k, mean(v2), sd(v2))
lines(k, gpdf, col = "red")
cat(sprintf("Skewness: %f\n", skewness(v2)))
cat(sprintf("Kurtosis: %f\n", kurtosis(v2)))
```
```{r}
#for B
hist(v3,breaks = 40,freq = FALSE)
k = seq(min(v3), max(v3), by = 0.001)
gpdf = dnorm(k, mean(v3), sd(v3))
lines(k, gpdf, col = "red")
cat(sprintf("Skewness: %f\n", skewness(v3)))
cat(sprintf("Kurtosis: %f\n", kurtosis(v3)))
```
9. Exploratory: Repeat steps 6-8. In this step assume that y=a+b*x+d where a=1.2, b=2.1, 
and d is from a t-distribution with degree of freedom as 3. 

```{r}
a=1.2
b = 2.1
x = seq(1, 5, by = 0.01)
y = numeric(length(x))
suppressWarnings({
  for (j in 1: length(x)){
  y[j] = a+b*x+rt(1,3)
}
})
train = as.data.frame(cbind(y,x))
mymod = lm(y~x, data = train)
ypred = predict(mymod, train)
me = mean(y-ypred)
ea = mymod$coefficients[1]
eb = mymod$coefficients[2]

```

```{r}
ntrials=1000
v1=numeric(ntrials)
v2=numeric(ntrials)
v3=numeric(ntrials)
for(i in 1:ntrials)
{
  a=1.2
  b=2.1
  x=seq(1,5,by=0.01)
  y=numeric(length(x))
  for(j in 1:length(x))
  {
    y[j]=a+b*x[j]+rt(1,3)#y=a+b(x)+d
  }
  #learn alinear regression model
  train=as.data.frame(cbind(x,y))
  mymod=lm(y~x,data=train)
  ypred=predict(mymod,train)
  v1[i]=mean(y-ypred) #variable 1
  v2[i]=mymod$coefficients[1] #variable 2
  v3[i]=mymod$coefficients[2] #variable 3
}
#for ME
hist(v1,breaks = 40,freq = FALSE)
k = seq(min(v1), max(v1), by = 1e-17)
gpdf = dnorm(k, mean(v1), sd(v1))
lines(k, gpdf, col = "red")
cat(sprintf("Skewness: %f\n", skewness(v1)))
cat(sprintf("Kurtosis: %f\n", kurtosis(v1)))
```
```{r}
#for A
hist(v2,breaks = 40,freq = FALSE)
k = seq(min(v2), max(v2), by = 0.001)
gpdf = dnorm(k, mean(v2), sd(v2))
lines(k, gpdf, col = "red")
cat(sprintf("Skewness: %f\n", skewness(v2)))
cat(sprintf("Kurtosis: %f\n", kurtosis(v2)))
```


```{r}
#for B
hist(v3,breaks = 40,freq = FALSE)
k = seq(min(v3), max(v3), by = 0.001)
gpdf = dnorm(k, mean(v3), sd(v3))
lines(k, gpdf, col = "red")
cat(sprintf("Skewness: %f\n", skewness(v3)))
cat(sprintf("Kurtosis: %f\n", kurtosis(v3)))
```
```{r}

```
