---
title: "experiment_6"
output: html_document
date: "2025-11-10"
---
1. Download the Boston dataset from library MASS. Find the column wise mean and standard deviation values. These are the true population parameters for this
experiment. 
```{r}
library(MASS)
data("Boston")
X = Boston
meanV = vector("numeric", ncol(X)) #to store the values of columnwise mean
sdV = numeric(ncol(X)) #to store the values of columnwise standard deviation

#Calculating mean using apply
meanV = apply(X, 2, mean)

#Calulcating standard deviation
sdV = apply(dataset, 2, sd)

meanV
sdV
```
2. One sample Mean Z-Test: Consider the rm column. 40 samples are randomly selected. A Null hypothesis is given to you that that true mean (u0) is equal to 6.2.
Based on the available samples, suggest whether you should accept this hypothesis or reject this hypothesis. Assume a significance level of 5%.

```{r}
rmCol =  X[,6]
s = sample(rmCol, size=40, replace = FALSE)
u0 = 6.2
alpha = 0.05
s_mean = mean(s) #sample statistic in this case
z = (s_mean-u0)/sdV[6]/sqrt(length(s)) #z-test assumes that true sd is more
p_value = 2*pnorm(abs(z), mean = 0, sd = 1, lower.tail = FALSE)
print(p_value)
if (p_value <0.05){
  print("Reject the null hypothesis")
}else{
  print("Accept the null hypothesis")
}
#
```

3. One sample proportion Z-Test: Consider the crim column. Create an indicator variable (0 or 1) high_crime. A location is high crime if the crime rate is greater than
median crime rate. We are interested in the proportion of locations which have a high crime rate. 40 samples (size=40) are randomly selected. A Null hypothesis is given to you
that that true proportion (p0) is equal to 0.5. Based on the available samples, suggest whether you should accept this hypothesis or reject this hypothesis. Assume that the
standard deviation of sample proportion is given by sqrt(p0*(1-p0)/size). Assume a significance level of 5%.

```{r}
p0 = 0.5
alpha = 0.05
size= 40
high_crime = (X$crim>median(X$crim))
s=sample(high_crime, size = size, replace = FALSE)
obs_proportion = mean(s) #statistic in this case
Z = (obs_proportion-p0)/sqrt(p0*(1-p0)/size) #sd for obs_proportion if null hypothesis is true
p_value = 2*pnorm(abs(Z), mean = 0 , sd = 1, lower.tail = FALSE)
if (p_value <0.05){
  print("Reject the null hypothesis")
}else{
  print("Accept the null hypothesis")
}
```

4. Two-Sample Difference of Means Z-Test: Consider the medv column. Split the column into two sets: locations where chas=1 and locations where chas=0. 20 samples
are randomly selected from each set. A Null hypothesis is given to you that that true mean of both sets are equal. Based on the available samples, suggest whether you should
accept this hypothesis or reject this hypothesis. Assume that standard deviation of difference of two sample means is given as total_sd=sqrt(sam1_sd^2+sam2_sd^2),
where sam1_sd and sam2_sd are the standard deviation of individual sample means. Assume a significance level of 5%.

```{r}
d0 = 0
alpha = 0.05
size = 20

s1 = X$medv[X$chas==1] #price of location near the charles river
s2 = X$medv[X$chas==0] #price of location away from the charles river

s1_tm = mean(s1) #true mean for set 1
s2_tm = mean(s2) #true mean for set 2

sam1 = sample(s1, size = 20, replace = FALSE)
sam2 = sample(s2, size = 20, replace = FALSE)

d_obs = mean(s1)-mean(s2) #statistic


sam1_sd = sd(sam1)/sqrt(size)
sam2_sd = sd(sam2)/sqrt(size)

total_sd=sqrt(sam1_sd^2+sam2_sd^2)


#statistic is 0 mean; sd = tot_sd; distribution = Gaussian

z = d_obs/total_sd #scaled statistic
p_value = 2*pnorm(abs(z), mean = 0 , sd = 1, lower.tail = FALSE)
p_value
if (p_value <0.05){
  print("Reject the null hypothesis")
}else{
  print("Accept the null hypothesis")
}

```

5. Two-Sample Difference of proportions Z-Test: Consider the tax column. Create an indicator variable (0 or 1) high_tax. A location is high tax if the tax is greater than median
tax. We are interested in the proportion of locations which have a high tax rate. Split the column into two sets: locations where chas=1 and locations where chas=0. 20 samples
are randomly selected from each set. A Null hypothesis is given to you that that true proportions of both sets are equal. Based on the available samples, suggest whether you
should accept this hypothesis or reject this hypothesis. Assume that standard deviation of sample proportion is sqrt(p(1-p)(2/size)) where p is the combined proportion of both sets.
Assume a significance level of 5%.

```{r}
dp0=0
alpha = 0.05
size = 20

high_tax = (X$tax>median(X$tax))

s1 = high_tax[X$chas==1] #price of location near the charles river
s2 = high_tax[X$chas==0] #price of location away from the charles river

tp_s1 = mean(s1)
tp_s2 = mean(s2)

sam1 = sample(s1, size, replace= FALSE)
sam2 = sample(s2, size, replace= FALSE)

obs_dp = mean(sam1) - mean(s2)
combined_p = mean(c(s1,s2))
obs_sd = sqrt(combined_p*(1-combined_p)/size)

Z = obs_dp-0/obs_sd
p_value = 2*pnorm(abs(z), mean = 0 , sd = 1, lower.tail = FALSE)
p_value
if (p_value <0.05){
  print("Reject the null hypothesis")
}else{
  print("Accept the null hypothesis")
}

```
6. Correlation Z-Test: Consider the rm and medv columns. Find the correlation using all samples. This is the true correlation between these two variables. Sample 40
observations randomly and test the Null Hypothesis that the two variables are uncorrelated. Use Fisher Transformation 0.5*log((1+r)/(1-r)) to convert r to normally
distributed random variable. The standard deviation of the obtained random variable is sqrt(1/(size-3)). Assume a significance level of 5%.

```{r}
r = 0
alpha = 0.05
size = 40
x1 = X$rm
x2 = X$medv
true_r = cor(x1, x2)
s1 = sample(x1, size, replace = FALSE)
s2 = sample(x2, size, replace = FALSE)

sam_r = cor(s1, s2)

#convert sam_r to Gaussian Distribution
z_r = 0.5*log((1+sam_r)/(1-sam_r))
sd_r = sqrt(1/(size-3))
z = z_r/sd_r
p_value = 2*pnorm(abs(z), mean = 0 , sd = 1, lower.tail = FALSE)
p_value
if (p_value <0.05){
  print("Reject the null hypothesis")
}else{
  print("Accept the null hypothesis")
}

```

7. Chi-square Independence Test: Create two indicator variables as high_crime and high_tax. Run chi-square independence test on these two variables and identify if these
variables are independent. A chi-square test assumes a Null Hypothesis that two variables are independent. Assume a significance level of 5%

```{r}
high_crime = (X$crim>median(X$crim))
high_tax = (X$tax>median(X$tax))
alpha = 0.05

test = chisq.test(high_crime, high_tax)
p_value = test$p.value
p_value
if (p_value <0.05){
  print("Reject the null hypothesis")
}else{
  print("Accept the null hypothesis")
}
```

8. Exploratory: Repeat step 2 for various Null Hypothesis: u0={5, 5.6, 6.2, 6.8, 7.4}. Record your p-values in each case.

```{r}
suppressWarnings({
  rmCol = X[,6]
  s = sample(rmCol, size = 40, replace = FALSE)
  u0 = c(5, 5.6, 6.2, 6.8, 7.4)
  alpha = 0.05
  s_mean = mean(s)
  p_value = numeric(length(u0))
  
  for (i in 1:length(u0)){
    z = (s_mean - u0[i]) / (sdV[6] / sqrt(length(s))) # standard error
    p_value[i] = 2 * pnorm(abs(z), lower.tail = FALSE)
    
    if (p_value[i] < alpha){
      print(paste("Reject the null hypothesis for u0 =", u0[i], "p_value = ", p_value[i]))
    } else {
      print(paste("Accept the null hypothesis for u0 =", u0[i], "p_value = ", p_value[i]))
    }
  }
})
```

9. Exploratory: Repeat step 3 for various Null Hypothesis: p0={0.2,0.4,0.6,0.8}. Record
your p-values in each case.

```{r}
p0 = c(0.2, 0.4, 0.6, 0.8)
alpha = 0.05
size = 40

high_crime = (X$crim > median(X$crim))
s = sample(high_crime, size = size, replace = FALSE)
obs_proportion = mean(s)

p_value = numeric(length(p0))

for (i in 1:length(p0)) {
  Z = (obs_proportion - p0[i]) / sqrt(p0[i] * (1 - p0[i]) / size)
  p_value[i] = 2 * pnorm(abs(Z), lower.tail = FALSE)
}
for(i in 1:length(p0)){
  if(p_value[i] < alpha){
    print(paste("Reject null hypothesis for p0 =", p0[i], "p_value: ", p_value[i]))
  } else {
    print(paste("Accept null hypothesis for p0 =", p0[i], "p_value: ", p_value[i]))
  }
}
```


